A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-10-28 imaging_demo/ms.tex
Make it compatible with arXiv (submitted v1)
From 70146f3 (arxiv):
====================

In ms.tex,

(*) Change including .bib file to including .bbl file
(*) Remove commented lines
(*) Remove figure filename extensions
(*) Comment out \usepackage{newtxtext,newtxmath} to be compatible
with arXiv submission
(*) Change fontsize of first table to scriptsize to fit the column
width

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{../mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

%%%%%%%%%

%We provide a brief summary of the mathematical equivalence of the MOFF and FX
%correlators detailed in \citet{mor11}. 
%
%Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky
%coordinate system denoted by unit vector $\hat{\mathbf{s}}$ propagate towards
%the observer as:
%\begin{align}
%  \widetilde{E}(\mathbf{r}) &= \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}},
%\end{align}
%where, $\mathbf{r}$ denotes the observer's coordinate system and $\widetilde{E}
%(\mathbf{r})$ is the propagated electric field. Thus the propagated electric
%field is a linear superposition of the electric fields emanating from
%astronomical sources with appropriate complex phases. It can also be described
%as a Fourier transform of the electric fields in the sky coordinates. 
%
%An antenna, $a$, measures a phased sum of these propagated electric fields over
%its effective collecting area with an additive receiver noise:
%\begin{align}\label{eqn:measured-E-field}
%  \widetilde{E}_a &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a)\,\widetilde{E}(\mathbf{r})\,\dif^2\mathbf{r} + \widetilde{n}_a \\
%                  &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \left[ \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a \\
%                  &= \int \delta(\mathbf{r}-\mathbf{r}_a) \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a
%\end{align}
%where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination
%pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$ is the
%directional antenna voltage response.
%
%Interferometers measure {\it visibility} -- the degree of coherence -- between
%electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
%Visibility, $\widetilde{V}_p$, can be written as:
%
%\begin{align}
%  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle \label{eqn:cc-vis}\\
%                  &= \int\delta(\mathbf{u}-\mathbf{u}_p)\left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,B(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right] \dif^2\mathbf{u} + n_p,
%\end{align}
%where, $I(\hat{\mathbf{s}})$ is the sky brightness, $B(\hat{\mathbf{s}})$ is the
%antenna power response, $\mathbf{u}$ denotes the coordinates of the $uv$-
%measurement plane, $\mathbf{u}_p\equiv \mathbf{r}_a-\mathbf{r}_b$ is the antenna
%spacing vector corresponding to the antenna pair $p$, and $n_p$ is the receiver
%noise. This signifies that the visibility ($\widetilde{V}_p$) measured between a
%pair of antennas ($p$) is obtained by the multiplying the sky brightness
%$I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and
%Fourier transform from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
%coordinates, which are then sampled at the locations of the antenna spacings (or
%baselines), namely, $\mathbf{u}_p$ using the sampling function
%$\delta(\mathbf{u}-\mathbf{u}_p)$ and added to the receiver noise $n_p$. This can
%be equivalently re-written as:
%\begin{align}\label{eqn:software-holography}
%  \widetilde{V}_p &= \int\delta(\mathbf{u}^\prime-\mathbf{u}_p)\left[\int \tilde{B}(\mathbf{u}^\prime-\mathbf{u}) \right. \nonumber\\
%  &\left. \qquad \times \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int}\right] \dif^2\mathbf{u}^\prime + n_p,
%\end{align}
%where, $\tilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
%obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
%multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
%convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space. This is the software
%holographic equivalent of traditional FX correlator output. Following the matrix
%notation of \citet{mor11}, the above equation can be expressed as:
%\begin{align}
%  \mathbf{m}(\mathbf{v}) &= \widetilde{B}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,I(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
%\end{align}
%where, the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using
%$\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence
%function is weighted and summed using the antenna power response,
%$\widetilde{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline
%location to obtain the measured visibilities:
%\begin{align}
%  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{E}(\mathbf{a})\ast\widetilde{E}(\mathbf{a}^\prime)\right\rangle_t. \label{eqn:matrix-cc-vis}
%\end{align}
%$\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating
%measured antenna electric fields over all possible pairs of $\mathbf{a}$ and
%$\mathbf{a}^\prime$. It is the same as equation~\ref{eqn:cc-vis} written in
%matrix notation.
%
%Using the optimal map-making formalism \citep{teg97a,teg97b}, a software
%holography image is formed using \citep{mor09}:
%\begin{align}
%  I^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
%\end{align}
%where the measured visibilities are weighted by the inverse of the system noise,
%following a gridding process using the holographic antenna power response as the
%gridding kernel followed by a Fourier transform to create an image
%$I^\prime(\hat{\mathbf{s}})$ which is the optimal estimate of the true image
%$I(\hat{\mathbf{s}})$.

%It may be noted that the antenna power response in $uv$-plane can be written as:
%\begin{align}\label{eqn:antenna-power-breakup}
%  \tilde{B}_p(\mathbf{u}) &= \widetilde{W}_a(\mathbf{u}) \ast \widetilde{W}_b^\star(\mathbf{u})
%\end{align}
%where, $W_a(\mathbf{u})$ and $W_b(\mathbf{u})$ denote the individual antenna
%voltage responses. Thus $\tilde{B}_p(\mathbf{u})$ is a cross-correlation of the
%individual antenna voltage responses. Similarly, it can be shown that 
%\begin{align}
%  \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime)\right\rangle &= \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s}}^\prime) \right\rangle \nonumber\\
%  &= \left\langle \left|E(\hat{\mathbf{s}})\right|^2 \right\rangle = I(\hat{\mathbf{s}}) \label{eqn:sky-intensity}
%\end{align}
%
%From equations~\ref{eqn:measured-E-field}, \ref{eqn:software-holography},
%\ref{eqn:matrix-cc-vis}, \ref{eqn:antenna-power-breakup} and
%\ref{eqn:sky-intensity}, the optimal image in equation \ref{eqn:dirty-image-FX}
%can be expressed equivalently as:
%\begin{align}
%  I^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}(\mathbf{a},\mathbf{a})\,\widetilde{E}(\mathbf{a})\,\right|^2\right\rangle_t. \label{eqn:dirty-image-MOFF}
%\end{align}
%The term inside the angular brackets before squaring has a very similar form as
%that in equation~\ref{eqn:dirty-image-FX}. It signifies that the measured antenna
%electric fields are weighted by the antenna noise, weighted and gridded by the
%antenna aperture kernel, Fourier transformed and finally squared to obtain the
%same image estimated that would have been obtained using equation~
%\ref{eqn:dirty-image-FX}.

%Equations~\ref{eqn:dirty-image-MOFF} and \ref{eqn:dirty-image-FX} are identical
%to each other obtained as a consequence of the multiplication-convolution theorem
%of Fourier transforms. In other words, equation~\ref{eqn:dirty-image-FX} is
%obtained by a Fourier transform of the gridded spatial correlation of electric
%fields measured by antennas while equation~\ref{eqn:dirty-image-MOFF} is obtained
%by squaring the Fourier transform of the gridded electric fields measured by the
%antennas. The time-scale of averaging in equations~\ref{eqn:cc-vis} and
%\ref{eqn:dirty-image-MOFF} is set by the coherence time-scale of the electric
%fields or the science requirement (e.g., timescale of variability or transients).
%
%There are some important differences between the two techniques:
%\begin{enumerate}
%\item The time-averaging cannot be performed on a stochastic measurement but
%  only on its statistical properties. In FX imaging, the visibilities measured
%  between antenna pairs represent spatial correlations which can be time-averaged
%  following which the they gridded and imaged. However, in MOFF imaging both
%  antenna and gridded electric fields are stochastic and therefore have to be
%  imaged before time-averaging. 
%\item In FX imaging, electric fields measured on antennas are not correlated with
%  themselves and hence lack zero spacing measurements. In contrast, in MOFF
%  imaging, since the gridded electric fields are imaged and squared, they
%  contain information from auto-correlated electric fields at zero spacing.
%  Hence, they have to be subtracted from the images.
%\end{enumerate}

@@

Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
propagated electric fields are shown on the left at different time stamps,
$t_1\ldots t_\textrm{M}$. At each time stamp, the electric fields measured by
antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$. The F-engine performs a
temporal Fourier transform on the electric field time-series to obtain electric
field spectra $E_1(f)\ldots E_\textrm{N}(f)$ 
($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix notation) for each of the 
antennas. Each of the complex antenna gains are calibrated to correct the 
corresponding electric field spectra. These calibrated electric fields are 
gridded using an antenna-based gridding convolution function after which it is 
spatially Fourier transformed and squared to obtain an image cube for every time 
stamp. These images are then time-averaged to obtain the accumulated image 
$I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
  \caption{A flowchart of MOFF imaging in EPIC. The propagated electric fields
    shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$
    by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then 
    Fourier transformed by the F-engine to produce electric field spectra 
    $E_1(f)\ldots E_\textrm{N}(f)$. They are calibrated and gridded. The gridded 
    electric fields $E_\textrm{g}(f)$ from each time series are imaged to produce 
    images $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to 
    obtain the final image $I(f)$.}
  \label{fig:MOFF-flowchart}
\end{figure}

Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility based software 
holographic imaging from a FX correlator. The antenna-based F-engine is identical 
to that in the MOFF processing. The electric field spectra from each antenna are 
then cross-multiplied in the X-engine with those from all other antennas to 
obtain the visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix 
notation). They are calibrated and time-averaged to obtain 
$\langle V_\textrm{ij}(f)\rangle$ which are then gridded and imaged to obtain the 
image $I(f)$. The $I(f)$ obtained from both techniques are theoretically 
identical as explained in \S\ref{sec:math}.
\begin{figure}
  \includegraphics[width=\columnwidth]{FX_flowchart.eps}
  \caption{A flowchart of visibility based software holographic imaging in EPIC. 
    The FX process flow shares the F-engine with the MOFF process. Following the 
    F-engine, the electric fields pass through the X-engine to produce 
    visibilities $V_\textrm{ij}(f)$ which are calibrated and time-averaged. Then 
    they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which 
    are then Fourier transformed to obtain the image, $I(f)$.}
  \label{fig:FX-flowchart}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{a2g_mapping.eps}
  \caption{Block diagram of an antenna-to-grid mapping. A sparse block-diagonal
    matrix of total size $\Ngrid\times \Nant$ is created where each
    block contains roughly the number of pixels covered by the respective kernel.
    The antenna aperture illumination kernels do not have to be identical to each
    other. A discrete set of arbitrarily placed antennas are now placed onto a
    regular grid.}
  \label{fig:a2g-mapping}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{F-engine.eps}
  \caption{Block diagram of a F-engine. The electric field data streams from
    antennas are Fourier transformed in parallel to generate electric field
    spectra.}
  \label{fig:f-engine}
\end{figure}

@@

\par\medskip
\noindent {\bf Gridding Convolution}
\par\medskip
\noindent The antenna array aperture illumination over the entire grid,
$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the 
individual antenna aperture illuminations:
\begin{align}\label{eqn:gridding-convolution}
  \widetilde{\mathbf{W}}(\mathbf{r}) &= \sum_a \widetilde{\mathbf{W}}_a(\mathbf{r}-\mathbf{r}_a) \\
                            &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathcal{I}(\mathbf{a}),
\end{align}
where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved by
efficient multiplication with the sparse matrix created in the antenna-to-grid
mapping process using the sparse matrices module in Python SciPy package. 
Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
effects of the ionosphere or the instrument, it needs to be computed just once
for the entire observation. However, the gridding of electric fields must be
computed at every readout of the electric field spectra,
\begin{align}
  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}).
\end{align}
% The sentence below was already stated earlier.
%In practice the grid is set to be a power of 2 such that the grid spacing is at
%most $\lambda/2$ even at the highest frequency in the observing band.

@@

\begin{figure*}
  \includegraphics[width=\linewidth]{EPIC-modules.eps}
  \caption{Software architecture of EPIC with core modules, their essential
    attributes and functions. The antenna module forms the fundamental building 
    block. It consists of electric field time-series and spectra and the 
    F-engine that performs a temporal FFT to obtain electric field spectra from 
    the time-series. The interferometer module is made of a pair of antenna 
    modules. Its main function is the X-engine (FX or XF) to produce visibility 
    spectra. The antenna array module is made of all individual antenna modules
    as its components and contains collective properties about the antenna 
    subsystems. Its core function is the creation of antenna-to-grid mapping, 
    gridded aperture weights and electric fields. The interferometer array 
    module is very similar in principle to the antenna array module except it 
    operates on cross-correlations and produces gridded visibilities. The image 
    module takes gridded electric fields or visibilities and performs a 
    two-dimensional spatial FFT (and squares the intermediate image in case 
    of the former) to produce output images. Broadly, the MOFF algorithm is 
    implemented by modules below the horizontal dashed line while the visibility 
    based imaging uses modules above the line. The exact processing pathway 
    implementing the MOFF algorithm is shown in bolded modules.}
  \label{fig:software-modules}
\end{figure*}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
  \caption{The auto-correlation of weights of a square shaped antenna aperture
    in the $uv$ plane (left) and the corresponding directional antenna power 
    response on the sky (right) in coordinates specified by direction cosines. 
    The antenna auto-correlation weights are normalized to a sum
    of unity yielding a peak response of unity in the antenna's directional
    power pattern on the sky. The color scale for the directional power 
    pattern is logarithmic. The black circle indicates the sky horizon and
    values beyond it are not physical and hence ignored.}
  \label{fig:autocorr_wts_PB}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
    simulated data using antenna-based MOFF algorithm (left) and FX 
    visibility-based software holography (right). The solid black circles in the
    top panels indicate the simulated source positions. The antenna 
    auto-correlations at zero-spacing have been removed from the MOFF images. 
    The images in either case reconstruct the sources at the right locations 
    with the fluxes expected after multiplication by the antenna power pattern. 
    The synthesized beams from the two algorithms are well matched in size and
    shape. The overall modulation by the power pattern is seen clearly in both
    images.}
  \label{fig:MOFF-FX-image}
\end{figure}

@@

% ----

% Fig.~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
% with MOFF imaging (left) and visibility based imaging (right). The first 
% notable difference is in the weights around zero spacing. Though both show a
% dominant void around zero-spacing, the void obtained with MOFF algorithm shows 
% many pixels with non-zero weights. In contrast, the zero-spacing void from 
% traditional imaging consists of predominantly zero-valued pixels. The gridding
% process in the former involves tying the antenna footprint (either in 
% analytic or lookup table formats) to the nearest grid pixel. Depending on the 
% exact location of the center of the antenna relative to the grid and the 
% coarseness of grid spacing, grid pixels that receive contribution from an 
% antenna may be a pixel narrower along one or both axes relative to that from 
% another identical antenna but with a different center location relative to the 
% grid. The resulting auto-correlation footprint will also not necessarily be 
% identical. Hence, using a single expected footprint for subtraction will 
% typically leave some residuals behind as seen in the void region in the weights 
% obtained with MOFF imaging.

% These residuals can be mitigated by: 
% \begin{enumerate}
%   \item making the grid spacing finer which makes the nearest neighbour 
%     association between the antenna footprint and the grid less susceptible 
%     (proportional to grid spacing) to the location of the antenna center 
%     relative to the grid, and
%   \item subtracting each auto-correlation of antenna weights separately by 
%     using the shape and extent of the footprint appropriate for that specific
%     antenna aperture. 
% \end{enumerate}
% The latter is the most general solution applicable especially in the case of 
% heterogeneous antenna arrays and is under active development for EPIC.

% The other notable difference is that weights outside the zero-spacing void in
% some regions are different from each other at the few percent level though the 
% sum of weights in these regions are identical. For instance, note the 
% difference in weights at $(u,v)\approx (2,-10)$. This is found to arise 
% because with MOFF imaging the antenna locations were tied to the nearest 
% grid pixel whereas in visibility based imaging the baseline locations 
% (difference between antenna locations) were tied to the nearest grid pixel. 
% Since rounding is not a commutative operation (i.e. rounding of the difference 
% value is not necessarily equal to the difference of the rounded values), the 
% gridding operation in the two cases introduces rounding errors in the placement 
% of antenna aperture and cross-correlated aperture weights. Each antenna 
% aperture weight projected to the nearest grid location can be displaced by 
% $\sim 1$--2 pixels. This effect can be mitigated only by making the grid 
% spacing finer at the expense of increased computational cost.

% ----

\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
  \caption{Comparison of $uv$ plane antenna power response with MOFF (left) and 
    visibility based imaging (right). The zero-spacing footprint from the $uv$ 
    weights obtained with MOFF has been subtracted. Differences in the void 
    centered around zero-spacing and other outer regions is due to loss of 
    sampled antenna footprint precision due to coarseness of grid spacing 
    further made severe by the correlation process.}
  \label{fig:MOFF-FX-uvwts}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
% \usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
% \usepackage{mathptmx}
% \usepackage{txfonts}

@@

@@

Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
propagated electric fields are shown on the left at different time stamps,
$t_1\ldots t_\textrm{M}$. At each time stamp, the electric fields measured by
antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$. The F-engine performs a
temporal Fourier transform on the electric field time-series to obtain electric
field spectra $E_1(f)\ldots E_\textrm{N}(f)$ 
($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix notation) for each of the 
antennas. Each of the complex antenna gains are calibrated to correct the 
corresponding electric field spectra. These calibrated electric fields are 
gridded using an antenna-based gridding convolution function after which it is 
spatially Fourier transformed and squared to obtain an image cube for every time 
stamp. These images are then time-averaged to obtain the accumulated image 
$I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_flowchart}
  % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
  \caption{A flowchart of MOFF imaging in EPIC. The propagated electric fields
    shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$
    by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then 
    Fourier transformed by the F-engine to produce electric field spectra 
    $E_1(f)\ldots E_\textrm{N}(f)$. They are calibrated and gridded. The gridded 
    electric fields $E_\textrm{g}(f)$ from each time series are imaged to produce 
    images $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to 
    obtain the final image $I(f)$.}
  \label{fig:MOFF-flowchart}
\end{figure}

Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility based software 
holographic imaging from a FX correlator. The antenna-based F-engine is identical 
to that in the MOFF processing. The electric field spectra from each antenna are 
then cross-multiplied in the X-engine with those from all other antennas to 
obtain the visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix 
notation). They are calibrated and time-averaged to obtain 
$\langle V_\textrm{ij}(f)\rangle$ which are then gridded and imaged to obtain the 
image $I(f)$. The $I(f)$ obtained from both techniques are theoretically 
identical as explained in \S\ref{sec:math}.
\begin{figure}
  \includegraphics[width=\columnwidth]{FX_flowchart}
  % \includegraphics[width=\columnwidth]{FX_flowchart.eps}
  \caption{A flowchart of visibility based software holographic imaging in EPIC. 
    The FX process flow shares the F-engine with the MOFF process. Following the 
    F-engine, the electric fields pass through the X-engine to produce 
    visibilities $V_\textrm{ij}(f)$ which are calibrated and time-averaged. Then 
    they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which 
    are then Fourier transformed to obtain the image, $I(f)$.}
  \label{fig:FX-flowchart}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{a2g_mapping}
  % \includegraphics[width=\columnwidth]{a2g_mapping.eps}
  \caption{Block diagram of an antenna-to-grid mapping. A sparse block-diagonal
    matrix of total size $\Ngrid\times \Nant$ is created where each
    block contains roughly the number of pixels covered by the respective kernel.
    The antenna aperture illumination kernels do not have to be identical to each
    other. A discrete set of arbitrarily placed antennas are now placed onto a
    regular grid.}
  \label{fig:a2g-mapping}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{F-engine}
  % \includegraphics[width=\columnwidth]{F-engine.eps}
  \caption{Block diagram of a F-engine. The electric field data streams from
    antennas are Fourier transformed in parallel to generate electric field
    spectra.}
  \label{fig:f-engine}
\end{figure}

@@

\par\medskip
\noindent {\bf Gridding Convolution}
\par\medskip
\noindent The antenna array aperture illumination over the entire grid,
$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the 
individual antenna aperture illuminations:
\begin{align}\label{eqn:gridding-convolution}
  \widetilde{\mathbf{W}}(\mathbf{r}) &= \sum_a \widetilde{\mathbf{W}}_a(\mathbf{r}-\mathbf{r}_a) \\
                            &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathcal{I}(\mathbf{a}),
\end{align}
where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved by
efficient multiplication with the sparse matrix created in the antenna-to-grid
mapping process using the sparse matrices module in Python SciPy package. 
Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
effects of the ionosphere or the instrument, it needs to be computed just once
for the entire observation. However, the gridding of electric fields must be
computed at every readout of the electric field spectra,
\begin{align}
  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}).
\end{align}

@@

\begin{figure*}
  \includegraphics[width=\linewidth]{EPIC-modules}
  % \includegraphics[width=\linewidth]{EPIC-modules.eps}
  \caption{Software architecture of EPIC with core modules, their essential
    attributes and functions. The antenna module forms the fundamental building 
    block. It consists of electric field time-series and spectra and the 
    F-engine that performs a temporal FFT to obtain electric field spectra from 
    the time-series. The interferometer module is made of a pair of antenna 
    modules. Its main function is the X-engine (FX or XF) to produce visibility 
    spectra. The antenna array module is made of all individual antenna modules
    as its components and contains collective properties about the antenna 
    subsystems. Its core function is the creation of antenna-to-grid mapping, 
    gridded aperture weights and electric fields. The interferometer array 
    module is very similar in principle to the antenna array module except it 
    operates on cross-correlations and produces gridded visibilities. The image 
    module takes gridded electric fields or visibilities and performs a 
    two-dimensional spatial FFT (and squares the intermediate image in case 
    of the former) to produce output images. Broadly, the MOFF algorithm is 
    implemented by modules below the horizontal dashed line while the visibility 
    based imaging uses modules above the line. The exact processing pathway 
    implementing the MOFF algorithm is shown in bolded modules.}
  \label{fig:software-modules}
\end{figure*}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
  \caption{The auto-correlation of weights of a square shaped antenna aperture
    in the $uv$ plane (left) and the corresponding directional antenna power 
    response on the sky (right) in coordinates specified by direction cosines. 
    The antenna auto-correlation weights are normalized to a sum
    of unity yielding a peak response of unity in the antenna's directional
    power pattern on the sky. The color scale for the directional power 
    pattern is logarithmic. The black circle indicates the sky horizon and
    values beyond it are not physical and hence ignored.}
  \label{fig:autocorr_wts_PB}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
    simulated data using antenna-based MOFF algorithm (left) and FX 
    visibility-based software holography (right). The solid black circles in the
    top panels indicate the simulated source positions. The antenna 
    auto-correlations at zero-spacing have been removed from the MOFF images. 
    The images in either case reconstruct the sources at the right locations 
    with the fluxes expected after multiplication by the antenna power pattern. 
    The synthesized beams from the two algorithms are well matched in size and
    shape. The overall modulation by the power pattern is seen clearly in both
    images.}
  \label{fig:MOFF-FX-image}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed}
  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
  \caption{Comparison of $uv$ plane antenna power response with MOFF (left) and 
    visibility based imaging (right). The zero-spacing footprint from the $uv$ 
    weights obtained with MOFF has been subtracted. Differences in the void 
    centered around zero-spacing and other outer regions is due to loss of 
    sampled antenna footprint precision due to coarseness of grid spacing 
    further made severe by the correlation process.}
  \label{fig:MOFF-FX-uvwts}
\end{figure}
